---
title: 'STAT 243 Final Project: Genetic Algorithm  Documentation'
author: "Kunal Desai, Fan Dong, James Duncan, Xin Shi"
date: "December 6, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How $\textit{select}$ Works
(discuss our general approach here)

## An Example

To test the algorithm, we randomly generate a dataset of 40 predictors and 500 observations, in which 6 are real predictors (named $real_{1} \sim real_{6}$) and the rest are pure noise variables($noi_{1} \sim noi_{34}$). The response variable, $y$, is thus given by the equation (coefficients are picked randomly): 

$y= \beta_{0}+\beta_{1} \times real_{1} + \beta_{2} \times real_{2} + \beta_{3} \times real_{3} + \beta_{4} \times real_{4} + \beta_{5} \times real_{5} + \beta_{6} \times real_{6}$

Code for data simulation:
```{r }
n <- 500
c <- 40
X <- matrix(rnorm(n * c), nrow = n)
beta <- c(88, 0.1, 123, 4563, 1.23, 20)
y <- X[ ,1:6] %*% beta
colnames(X) <- c(paste("real", 1:6, sep = ""),
                     paste("noi", 1:34, sep = ""))

```

Ideally, our genetic algorithm will successfully select $real_{1} \sim real_{6}$ out of the 40 predictors. 

## How the Team Works

The project resides in Kunal's repository (Git Username: kunaljaydesai, Repo name: GA).

The specific tasks completed by each group member is listed below:
\begin{enumerate}
 \Kunal
 \Fan
 Wrote ranked_model, generation_gap and wrote their respective tests. Modified and finalized calculated_fitness and wrote its respective tests. Wrote the \textit{An Example} section in this documentation.
 \James
 \Xin
\end{enumerate}
